\section{What is this thesis about?}

Optimal transport theory has found many application in diverse fields in machine learning thank to 
providing a powerful tool for comparing probability distributions, one of a crucial issue
in machine learning.  In this thesis, we leverage the optimal transport theory and statistics to deal with 
the problems in biology and actuary. The biological problem is to assess the possible relationships between microRNA and mRNA expression in the striatum of Huntington's disease model mice. The actuarial problem relates to anticipate the declaration of natural disaster for a drought event. 

\paragraph{ Huntington’s disease\\}
Huntington's disease, an autosomal-dominant, progressive neurodegenerative disorder, is characterized by involuntary chromatic movements with cognitive and behavioral disturbances and is caused by an expansion of a repeating CAG triplet series in the huntingtin gene~\cite{HD2007, MACDONALD1993971}.   In normal individuals the CAG repeat length ranges from 10 to 35, while in HD individuals, it ranges from 36 to more than 120 which is inversely correlated with age of onset. Specifically, HD patients with 36-40 CAG repeats may have late onset or may not develop symptoms while repeat lengths in the 40s have symptom onset in the fourth decade and repeat lengths greater than 60 lead to juvenile onset \cite{Andrew1993}
There are currently no treatments to prevent the onset or slow the progression of HD.

Huntington's disease, like several neurodegenerative diseases such as Alzheimer's disease, Parkinson's disease and amyotrophic lateral sclerosis, relates to gene deregulation which has encouraged large studies to gene regulatory mechanisms~\cite{Langfelder2018}.
Gene expression is controlled by limiting the amount of mRNA produced from a particular gene at the transcription level and regulating the translation of mRNA into proteins at the post-transcriptional level. The most important instruments in the latter level  are small non-coding RNAs called miRNAs. It binds to a complementary sequence in the 3'UTR of the target mRNA resulting in a rapid degradation of the mRNA or less frequently in an inhibition of its translation into protein~\cite{Pasquinelli12}. So the researchers are interested in studying the interaction between miRNAs and mRNAs in HD to gain a deeper understanding the disease and to develop a new treatment.

\paragraph{Anticipate the declaration of natural disaster for a drought event in France}




\section{Formalisation}

\paragraph{Huntington’s disease}
In recent years, advanced sequencing technologies such as RNAseq produce large data sets as genome, proteome, transcriptome and metabolome, which are analysed to gain insight into genetics, human biology and disease. Notably, the data of Huntington's disease are increasingly available for the analyse of  molecular pathology such as  mRNA expression signatures of HD in post-mortem human brains~\cite{Neueder2014, CHA2007228} and the changes in miRNA expression  in multiple mouse models~\cite{Langfelder2018}.  However, little is known about the interaction between mRNA and miRNA in HD because of (1) lack of time-series and high quality-data from different cells/ tissues from healthy and diseased conditions; (2)difficulty to build an accurately modeling between miRNAs and mRNAs with an acceptable of false positive and false negative results (3) costly to carry out experimental methods and validation of miRNA target gene.

Encouraged by the promising findings of~\cite{megret:inserm-02512089, Langfelder2018}, our goal is to shed light on the interaction between mRNA and miRNAs based on multidimensional data which are collected at three different ages in striatum (a brain region) from an allelic series of HD model knock-in mice with increasing CAG length in the endogenous Huntingtin gene.  For each combination of poly Q length and age, we have quantified miRNA and mRNA expression of 8 mice including 4 females and 4 males. After preprocessing ~\cite{megret:inserm-02512089}, we obtained the final dataset consisting of $M= 13 616$ mRNA profiles and $N=1143$ miRNA profiles. The fact that miRNA and target mRNAs are linked by a "many-to-many" mirroring relationship because a miRNA induces the degradation of a target mRNA or blocks its translation into proteins, or both and a miRNA can regulate several mRNAs. The biological hypothesis guides to identify the negatively correlative miRNA-mRNA pairs.

\section{State of the art}

\paragraph{Hungtinton disease}

From the first discovery of miRNAs in 1993~\cite{LEE1993843}, numerous methods of computational prediction, experimental detection and validation of miRNA target genes have been studied  for understanding miRNA function in varied biological processes. Experimental methods are time-consuming and expensive especially with more than one miRNA of interest furthermore some experimental methods following overexpression of miRNAs were criticized for generating false positive results. Computational methods employ different stati





  
\paragraph{Anticipate declaration of natural disaster for a drought event}


\section{Design, programming and implementation of algorithms}
\section{Results}
